---
title: "Read, write and plot TIFFs"
author: "Volker J. Schmid, Bioimaging group, Department of Statistics, LMU Munich"
date: '`r Sys.Date()`'
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Read, write and plot}
  %\VignetteEncoding{UTF-8}
output:
  knitr:::html_vignette:
    toc: yes
---
```{r loadlibrary, cache=TRUE}
library(bioimagetools)
```

## Read and write 3D RGB tiff stacks

`readTIF()` and `writeTIF()` read and write tiff stacks. 

3D and color stacks are automatically recognized by `readTIF()`, if the meta information is correct. Available meta information is saved as attributes. `readTIF()` can also handle URLs. The function returns an array with range 0 to 1.
```{r readtif, warnings=FALSE, collapse=TRUE}
cell <- bioimagetools::readTIF("http://ex.volkerschmid.de/cell.tif")
print(dim(cell))
print(attributes(cell)$dim)
print(attributes(cell)$bits.per.sample)
par(pty="s")
bioimagetools::img(cell, z=25 ,col="rgb")
```

`writeTIF()` writes out tiff stacks. Meta information in attributes are saved. The bits per sample can be controlled by an option.
```{r writetif, warnings=FALSE, collapse=TRUE, cache=TRUE}
writeTIF(cell, file="my_cell.tif")
```

`readClassTIF()` is a special version of `readTIF()` for images with a very low number of intensity values, e.g. masks or segmented images.
```{r readclasstif, warnings=FALSE, collapse=TRUE, cache=TRUE}
bioimagetools::writeTIF(2*EBImage::thresh(cell[,,1,])+EBImage::thresh(cell[,,2,]), file="simple.tif")
mysimple <- readClassTIF("simple.tif", n=3)
par(pty="s")
bioimagetools::img(mysimple[,,25],col="red",up=3)
```
```{r remove, echo=FALSE}
file.remove("my_cell.tif")
file.remove("simple.tif")
```

## Plotting
`img()` is a plotting function for images stored in an array.  
```{r img, warnings=FALSE, collapse=TRUE, cache=TRUE}
par(pty="s")
bioimagetools::img(cell, z=25, col="rgb")
bioimagetools::img(cell, z=25, col="r")
bioimagetools::img(cell, z=25, col="grey")
```

## Reading Bitmap files

Simple Bitmap files can be imported by `readBMP()`.
```{r readMBP, warnings=FALSE, collapse=TRUE, cache=TRUE}
bi<-readBMP(system.file(package = "bioimagetools","extdata", "V.bmp"))
par(pty="s")
image(bi, col=grey(seq(0,1,length=2^8)))
```

## Single TIF slices

In order to work with single slices, you would like to select the slice from the array:

```{r singleslice1, warnings=FALSE, collapse=TRUE, cache=TRUE}
slice30<-cell[,,,30]
```

However, this will result in the wrong dimension. bioimagetools always needs 4d array:

```{r singleslicedim, warnings=FALSE, collapse=TRUE, cache=TRUE}
dim(slice30) # only 3d
dim(slice30) <- c(dim(slice30), 1)
dim(slice30) # 4d 
img(slice30, z=1, col="rgb")
writeTIF(slice30,paste0("slice30.tif"))
```

To read single RGB slices, you can use readTIFF. Make sure to add the option channels=3 for an RGB slice.

```{r readsingle, warnings=FALSE, collapse=TRUE, cache=TRUE}
slice30<-readTIF("slice30.tif", channels=3)
dim(slice30)
```

To stack single slices use abind from the abind package:

```{r abind, warnings=FALSE, collapse=TRUE, cache=TRUE}
for (i in 1:52)writeTIF(cell[,,,i],paste0("slice",i,".tif")) # write slices
slice<-list()
for (i in 1:52)slice[[i]]<-readTIF(paste0("slice",i,".tif"), channels=3) # read slices
allslices<-slice[[1]] # slice 1
for (i in 2:52)allslices<-abind::abind(allslices, slice[[i]]) # stack slices 2 to 52
dim(allslices)
```

The attributes will be off for the stacked array, so we need to add them from hand using one of the slices (not the dimension!):

```{r attributes, warnings=FALSE, collapse=TRUE, cache=TRUE}
attrib<-attributes(slice[[1]])
attrib<-attrib[names(attrib)!="dim"] # Remove dimension attribute 
attributes(allslices)<-c(attributes(allslices),attrib)
```

```{r cleanup, echo=FALSE, warnings=FALSE, collapse=TRUE, cache=TRUE}

for (i in 1:52)fs::file_delete(paste0("slice",i,".tif"))
```





